---
title: "Chapter5.Rmd"
author: "Alex Suomalainen"
date: "26 11 2020"
output: html_document
---
# Chapter 5 Analysis

Loading the data
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)
```



Assessing libraries.
```{r}
library(GGally)
library(tidyverse)
library(dplyr)
```

Overview of the data.
```{r}
summary(human)
str(human)
ggpairs(human)
```

The data has 155 obsevations of 8 variables.  
There are integrer and numerical variables.  

Performing PCA with the non scaled data  
drawing a biplot of the principal component representation and the original variables
```{r}
pca_human <- prcomp(human)
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

Creating a summary of pca_human
```{r}
s <- summary(pca_human)
s
```

This makes no sense whatsoever. 

Standardising variables and doing the same operations.
```{r}
human_std <- scale(human)
pca_human <- prcomp(human_std)
biplot(pca_human, choices = 1:2, cex = c(0.5, 0.3), col = c("grey40", "deeppink2"))
s <- summary(pca_human)
s
```

Labo.FM and Parli.F variables are correlated, which means that when there is more females with at least secondary education,  
there are more female representatives in the parlament.

Mat.Mor and Ado.Birth variables are correlated, which means that when there is a higher maternal mortality ratio,  
there are more adolescents giving birth. 

Edu.Exp, Edu2.M, GNI and Life.Exp are correlated. So life expectancy at birth, Gross National Income per capita,  
proportion of males with at least secondary education and Expected years of schooling are correlated.

PC1 and PC2 contribute to (0.5361 + 0.1624) * 100% = 69,24% of the total variance.











