---
title: "chapter4.Rmd"
author: "Alex Suomalainen"
date: "19 11 2020"
output: html_document
---

## Analysis chr4

Assessing libraries
```{r}
library(MASS)
library(tidyr)
library(readr)
library(ggplot2)
```

Loading Boston data
```{r}
data("Boston")
```

Exploring dataset
```{r}
str(Boston)
summary(Boston)
```
The dataset has 506 observations/rows and 14 variables/columns.  


This data frame contains the following columns:  
**crim** = per capita crime rate by town.  
**zn** = proportion of residential land zoned for lots over 25,000 sq.ft.  
**indus** = proportion of non-retail business acres per town.  
**chas** = Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).  
**nox** = nitrogen oxides concentration (parts per 10 million). rm = average number of rooms per dwelling.  
**age** = proportion of owner-occupied units built prior to 1940.  
**dis** = weighted mean of distances to five Boston employment centres.  
**rad** = index of accessibility to radial highways.  
**tax** = full-value property-tax rate per \$10,000.  
**ptratio** = pupil-teacher ratio by town.  
**black** = 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.  
**lstat** = lower status of the population (percent).    
**medv** = median value of owner-occupied homes in \$1000s.

Most of the variables are numerical variables, with a couple of integrer variables.  

Graphical overview of the data
```{r}
pairs(Boston)
```


Scaling the dataset.  
```{r}
boston_scaled <- scale(Boston)
summary(boston_scaled)
```
The variables changes so that all the means are 0.  

Finding the class
```{r}
class(boston_scaled)
```
The class of boston scaled is matrix  

Creating a data frame  
```{r}
boston_scaled <- as.data.frame(boston_scaled)
class(boston_scaled)
```
Now it's a data frame. 

Creating quantile vector of crime rate  
```{r}
bins <- quantile(boston_scaled$crim)
bins
```

Creating a categorial variable 'crime'
```{r}
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, label = c( "low", "med_low", "med_high", "high"))
table(crime)
```
The categories are evenly distributed. 

Removing the original variable  
and adding the new variable
```{r}
boston_scaled <- dplyr::select(boston_scaled, -crim)
boston_scaled <- data.frame(boston_scaled, crime)

```

Dividing the data to train and test sets with 80% of the data in the train set
```{r}
n <- nrow(boston_scaled)
ind <- sample(n,  size = n * 0.8)
train <- boston_scaled[ind,]
test <- boston_scaled[-ind,]
```

Taking the crime classes from the test and saving them as correct_classes  
Executing the code to remove crime from test set
```{r}
correct_classes <- test$crime
test <- dplyr::select(test, -crime)
```

Fitting the linear discriminant analysis on the train.  Target variable is crime rate and all other variables are predictor variables
```{r}
lda.fit <- lda(crime ~ ., data = train)
lda.fit
```

Adding biplot arrows, changing the classes to numeric and drawing the plot.    
```{r}
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "orange", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}
classes <- as.numeric(train$crime)
plot(lda.fit, dimen = 2, col = classes, pch = classes)
lda.arrows(lda.fit, myscale = 1)
  
```

Predicting the classes with the test data
```{r}
lda.pred <- predict(lda.fit, newdata = test)
table(correct = correct_classes, predicted = lda.pred$class)
```

The extremes were predicted pretty correctly  
There is some difference in the medium classes.  




