---
title: "chapter3.rmd"
author: "Alex Suomalainen"
date: "12 11 2020"
output: html_document
---

---
title: "chapter3.rmd"
author: "Alex Suomalainen"
date: "12 11 2020"
output: html_document
---

# Regression and model validation
## Analyze

reading the data into r
```{r}
alc <- read.csv('alc.csv', TRUE, ",")
```

cheking the names of the variables
```{r}
str(alc)
```
This data has 36 variables and I'm choosing age, go out, studytime and activities for the analyze.  
age - student's age (numeric: from 15 to 22)  
goout - going out with friends (numeric: from 1 - very low to 5 - very high)   
studytime - weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)  
traveltime - home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)

Hypothesis:  
age ↑  alcohol consunption ↑  
going out with friends ↑ alcohol consumption ↓  
studytime ↑ alcohol consumption ↓  
traveltime ↑ alcohol consumption ↑

assessing libraries 
```{r}
library(tidyverse); library(tidyr); library(ggplot2); library(dplyr)
```

first I look at age and alcohol comsumption
```{r}
ggplot(alc, aes(x = high_use , y = age  )) +
  geom_boxplot()
```

High usage of alcohol seems to rise as age rises.  

Next lets take a look at go out time
```{r}
ggplot(alc, aes(x = high_use, y = goout  )) +
  geom_boxplot()
```

Alcohol usage seems to increase as go out time increases, which makes sense.  

Next up, studytime vs. alcohol consumption
```{r}
ggplot(alc, aes(x = high_use, y = studytime  )) +
  geom_boxplot()
```

With more study time there is less high alcocol usage. 

Last up, traveltime. This should be interesting
```{r}
ggplot(alc, aes(x = high_use, y = traveltime  )) +
  geom_boxplot()
```

There seems to be no correlation with traveltime and high alcohol usage

## Logistic regression

find the model with glm()  
```{r}
m <- glm( high_use ~ alc$age + alc$goout + alc$studytime + alc$traveltime, data = alc, family = "binomial" )
```

summarising
```{r}
summary(m)
```

All variables except age have a relatively low p-value suggesting that there is a relationship betweed them and high alcohol comsumption  

Lets compute the odds ratios  
```{r}
OR <- coef(m) %>% exp
```
and confidence intervals 
```{r}
CI <- confint(m) %>% exp
```

Then lets print out the odds ratios with the intervals
```{r}
cbind(OR, CI)
```

It seems that most likely when go out time increases, high alcohol consumption increases.  
It also seems like the more you study the less likely it is that you comsume a lot of alcohol. 
These were the correct hypotheses.

I don't see a strong correlation with age and high alcohol usage, since the OR is only 1.12.

Thats all I'm able to do. This math stuff is all new to me so this could be totally wrong. 

